Reasoning About a Machine with Local Capabilities

### 2018-04-27 toplas/paper.tex
Added some text for the malloc specification section

\begin{toplas}
  \section{Malloc and macros}
  \label{sec:malloc-and-macros}
  Assembly code can be difficult for people to easily read and understand as it is close to byte code and offers none of the abstractions of high-level languages. In \sectionname~\ref{sec:examples}, we present a number of program examples for the simple capability machine we present in this paper. In order to make these examples readable, we introduce a series of macros that each corresponds to a series of actual instructions on the machine.

  In \sectionname~\ref{sec:examples}, we will prove correctness lemmas for each of the program examples. Some of the program examples use malloc and with the logical relation defined in \sectionname~\ref{sec:logical-relation}, we can present the specification of malloc.

  In this section, we present the specification of malloc as well as the macros used in \sectionname~\ref{sec:examples}.

\subsection{Malloc specification}
  \itoplassug{Suggestion: add the malloc specification}
\begin{definition}[Malloc Specification]
  \label{spec:malloc}
  $c_\malloc$ satisfies the specification for malloc iff
  \[  
    \begin{array}{l}
      c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
      \exists \iota_{\malloc,0} \ldotp \\
      \quad \left(\forall \iota' \futurestr \iota_{\malloc,0} \ldotp \forall W,i \ldotp W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)])) \right) \wedge \\
      \quad \iota_{\malloc,0}.v = \perma \; \land \\
      \quad (\forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
      \qquad \forall i, n, \size \in \nats \ldotp \forall
      w_{\var{ret}} \in \Words \ldotp \\
      \qquad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
      \qquad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
      \qquad \quad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
      \qquad \quad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
      \qquad \quad \Rightarrow \\
      \qquad \qquad\exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
      \qquad \qquad \quad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
      \qquad \qquad \qquad \Phi \step[j] \Phi' \land \\
      \qquad \qquad \qquad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
      \qquad \qquad \qquad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
      \qquad \qquad \qquad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
      \qquad \qquad \qquad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
      \qquad \qquad \qquad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
      \qquad \qquad \qquad \size - 1 = e'-b' ) \land\\
      \quad (\forall \Phi \in \ExecConfs \ldotp (\memreg(r_1)
      \not\in \ints \vee \memreg(r_1) < 0) \land \memreg(\pcreg) = \updatePcPerm{c_\malloc} \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed)
    \end{array}
  \]
\end{definition}

>>>>

\begin{toplas}
  \section{Malloc and macros}
  \label{sec:malloc-and-macros}
With the semantic model defined, we would like to show its usefulness by proving
the correctness of a series of examples that use the calling convention and rely
on well-bracketedness and local-state encapsulation. However, the programs that
run on our capability are assembly programs and even program examples that would
be small in a high-level language, they become big and unintelligible at this
low level. In order to make our program examples intelligible, we introduce a
series of low-level abstractions in the form of macros.

In order to make the program examples realistic, we allow them to allocate fresh
memory by invoking a trusted malloc. Our results are not specific to
any one implementation of malloc, so rather than providing one implementation,
we provide a specification that malloc must satisfy.

\subsection{Malloc specification}
  \itoplassug{Suggestion: add the malloc specification}
Before we give the malloc specification, we provide some intuition about what
the malloc specification should look like.

Malloc is assumed to be part of the trusted computing base, as mentioned in
\sectionname~\ref{sec:capab-mach-with}. A program with access to malloc should
be able to retain this access even if it invokes code it does not trust. In
order to achieve this, a capability for malloc must be global. It is also
important that malloc remains trustworthy to all entities on the machine
throughout execution. That is, malloc should encapsulate its local state and
not depend on any external code. To make sure that malloc cannot be tampered
with through the capability for invoking malloc, we also require the capability
to have enter permission.

The requirements so far prevent direct tampering in the sense that the
capability intended for invoking malloc only can be used for invocation. This
does, however, not prevent the implementation of malloc from colluding with
other code. Specifically, malloc could depend on the state of a piece of memory
not just under the control of malloc. We also need to make sure that malloc
actually behave as intended. That is when malloc is invoked with a length
argument in the correct register, then it should actually return a capability
for a fresh piece of memory of that size. It should be fresh in the sense that
malloc hasn't given out a capability for any part of that memory before and
won't do it in the future. Finally, we also need to make sure that malloc
behaves correctly when given a nonsensical length argument such as a negative
integer or a capability. These requirements relate to the behaviour of malloc,
so we cannot hope to specify them syntactically. Instead we will use the
semantic model defined in  Section~\ref{sec:logical-relation} to express this
part of the specification.

In order to express the semantic requirements on malloc, the specification will
require the existence of some initial region for malloc. Malloc does not
disappear from the system, so the malloc region must permanent. This also means
that the malloc specification should guarantee that use of malloc is safe at any
point of execution and no matter who has invoked it. To this end, the
requirement on the initial malloc region should also hold true in any private
future world.

The implementation of malloc should not depend on any memory outside the
implementation of malloc. This means that malloc should not have any
capabilities in its internals that govern other programs. The malloc region's
memory interpretation function determines what malloc may contain, so we require
it to allow the same memory segments whether it is with respect to some world
with the malloc region in it or a world that just contains the malloc region.
Malloc should retain this property throughout execution, so it must hold true
for any private future region of the initial malloc region.

The malloc specification also dictates what malloc should do when invoked. We
specify the two cases where malloc is invoked with a valid and an invalid length
argument, respectively. When malloc is invoked with an invalid length argument,
that is a negative integer or a capability, we require malloc to simply fail.
This part of the specification does not rely on the malloc region as we do not
need to make sure that the memory invariants are respected. 
On the other hand,
\begin{definition}[Malloc Specification]
  \label{spec:malloc}
  $c_\malloc$ satisfies the specification for malloc iff
  \[  
    \begin{array}{l}
      c_\malloc = ((\entry,\glob),\_,\_,\_) \land\\
      \exists \iota_{\malloc,0} \ldotp \\
      \quad \iota_{\malloc,0}.v = \perma \; \land \\
      \quad \left(
        \begin{array}{l}
          \forall \iota' \futurestr \iota_{\malloc,0}, W,i \ldotp\\
          \quad W(i)=\iota' \Rightarrow \iota'.H (\iota'.s) (\xi^{-1}(W)) = \iota'.H (\iota'.s) (\xi^{-1}([i \mapsto W(i)]))
        \end{array}
        \right) \wedge \\
      \quad \left(
      \begin{array}{l}
        \forall \Phi \in \ExecConfs \ldotp \forall \ms_{\var{footprint}}, \hsframe \in \HeapSegments \ldotp \\
        \quad \forall i, n, \size \in \nats \ldotp \forall
        w_{\var{ret}} \in \Words \ldotp \\
        \quad \quad \forall \iota_\malloc \futurestr \iota_{\malloc,0} \land \\
        \quad \quad \memheap = \ms_{\var{footprint}} \uplus \hsframe \land \heapSat[\ms_{\var{footprint}}]{n}{[i \mapsto \iota_\malloc]} \land \\
        \qquad \memreg(r_1) = \size \land \size \geq 0 \land  \memreg(r_0) = w_{\var{ret}} \land \\
        \qquad \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\ %and this needs to change with ABI.
        \qquad \Rightarrow \\
        \qquad \quad \exists \Phi' \in \ExecConfs \ldotp \exists \ms_{\var{footprint}}', \ms_{\var{alloc}} \in \HeapSegments\ldotp\\
        \qquad \qquad \exists j \in \nats \ldotp j > 0 \land \exists b',e'\in \Addrs \ldotp \exists \iota_\malloc' \in \Regions \ldotp \\
        \qquad \qquad \quad \Phi \step[j] \Phi' \land \\
        \qquad \qquad \quad \memheap[\Phi']=\ms_{\var{footprint}}' \uplus \hs_{\var{alloc}} \uplus \hsframe \land\\
        \qquad \qquad \quad \iota_{\malloc}' \futurewk \iota_\malloc \land \\
        \qquad \qquad \quad \heapSat[\ms_{\var{footprint}}']{n-j}{[i \mapsto \iota_\malloc']} \land \\
        \qquad \qquad \quad \dom(\hs_{\var{alloc}}) = [b',e'] \land \forall a \in [b',e']\ldotp \hs_{\var{alloc}}(a) = 0  \land \\
        \qquad \qquad \quad \memreg[\Phi'] = \memreg[\Phi]\update{\pcreg}{\updatePcPerm{w_{\var{ret}}}}\update{r_1}{((\rwx,\glob),b',e',b')} \land \\
        \qquad \qquad \quad \size - 1 = e'-b'
      \end{array}
      \right) \land\\
      \quad \left(
      \begin{array}{l}
        \forall \Phi \in \ExecConfs \ldotp\\
        \quad(\memreg(r_1) \notin \ints \vee \memreg(r_1) < 0) \land
        \memreg(\pcreg) = \updatePcPerm{c_\malloc} \\
        \qquad \Rightarrow \exists j \in \nats \ldotp \Phi \step[j] \failed
      \end{array}
        \right)
    \end{array}
  \]
\end{definition}

\itoplassug{Write something about initial capabilities to malloc. We will always
assume that capabilities for malloc satisfy the malloc specification. The malloc
specification prevents malloc from giving away internal capabilities and the
capability machine semantics does not allow capabilities to appear out of thin air.}

\begin{lemma}[Malloc is safe to pass to adversary]
  \label{lem:malloc-in-vr}
  If $W(r) \futurestr \iota_{\malloc,0}$, then
  $\npair{c_\malloc} \in \stdvr(W)$ for all $n$.
\end{lemma}

\itoplassug{Add the assumption that the malloc capability in examples
  satisfies specification.}
